---
title: "A4: Full Assignment Template"
author: "YOUR NAME HERE"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, class.source="bg-success")
```

```{r message=FALSE, warning=FALSE}
# load packages

```


## Exericse 1

### Import `enrollment` and `survey` datasets

```{r include=FALSE}
# Import enrollment and survey datasets

```

<br>

### Joining the `enrollment` and `survey` data so that you have a list of the students in the class with their survey responses.

```{r}

```

<br>

### Using a join function to provide a list of students that did not fill out the survey.

```{r}

```

<br>

### Using a join function to provide a list of students that filled out the survey but have dropped the course.

```{r}

```

<br>


## Exercise 2

### Import `purchases` and `prices` datasets

```{r include=FALSE}
# Import purchases and prices datasets

```

### Joining the `purchases` and `prices` datasets to obtain a list purchases with their prices. 

```{r}

```

<br>

### Using the join functions and the `summarize` function to obtain an amount of total revenue for purchases. 

```{r}

```

<br>

### The total revenue per customer

```{r}

```


*** 

## Exercise 3

### Use some of the skills learned in Module A4 to create a line graph for 4 different carriers of Arrival Delays over 12 months.

```{r}
# Your code for Exercise 3 here




```



