---
title: "A4: Full Assignment Template"
author: "Saransh Rakshak"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, class.source="bg-success")
```

```{r message=FALSE, warning=FALSE}
# load packages
library(tidyverse)
library(dplyr)
library(nycflights13)
library(ggplot2)
```

## Exericse 1

### Import `enrollment` and `survey` datasets

```{r include=FALSE}
# Import enrollment and survey datasets
enrollment <- read.csv("enrollment.csv")
survey <- read.csv("survey.csv")
```

```{R}
glimpse(enrollment)
glimpse(survey)
```

<br>

### Joining the `enrollment` and `survey` data so that you have a list of the students in the class with their survey responses.

```{r}
student_surveys <- enrollment %>%
  left_join(survey, by = "id")

student_surveys
```

<br>

### Using a join function to provide a list of students that did not fill out the survey.

```{r}
no_survey <- left_join(enrollment, survey, by = "id") %>%
  filter(is.na(username)) %>%
  select(id, name.x) %>%
  rename(name = name.x)

# View the result
no_survey
```

<br>

### Using a join function to provide a list of students that filled out the survey but have dropped the course.

```{r}
dropped_surveyed <- right_join(enrollment, survey, by = "id") %>%
  filter(is.na(name.x)) %>%
  select(id, name.y, username) %>%
  rename(name = name.y)

dropped_surveyed
```

<br>


## Exercise 2

### Import `purchases` and `prices` datasets

```{r include=FALSE}
# Import purchases and prices datasets
purchases <- read.csv("purchases.csv")
prices <- read.csv("prices.csv")
```

```{R}
glimpse(purchases)
glimpse(prices)
```

### Joining the `purchases` and `prices` datasets to obtain a list purchases with their prices. 

```{r}
purchase_prices <- left_join(purchases, prices, by = "item")

purchase_prices
```

<br>

### Using the join functions and the `summarize` function to obtain an amount of total revenue for purchases. 

```{r}
total_revenue <- purchase_prices %>%
  summarise(total_revenue = sum(price, na.rm = TRUE))

total_revenue
```

<br>

### The total revenue per customer

```{r}
cust_total_revenue <- purchase_prices %>%
  group_by(customer_id) %>%
  summarise(total_revenue = sum(price, na.rm = TRUE))

cust_total_revenue
```


*** 

## Exercise 3

### Use some of the skills learned in Module A4 to create a line graph for 4 different carriers of Arrival Delays over 12 months.

```{r}
# Your code for Exercise 3 here

flights_summary <- flights %>%
  filter(carrier %in% c("AA", "DL", "UA", "WN")) %>%
  group_by(carrier, month) %>%
  summarise(
    count = n(),
    mean_arrival_delay = mean(arr_delay, na.rm = TRUE)
    ) %>%
  mutate(month = factor(month.name[month], levels = month.name))

ggplot(data = flights_summary, aes(x = month, y = mean_arrival_delay, group = carrier)) +
  geom_line(aes(color = carrier)) +
  geom_point(aes(color = carrier)) +
  labs(
    title = "Comparison of Arrival Delays over months for 4 carriers",
    subtitle = "AA (American Airlines), DL (Delta), UA (United), WN (Southwest)",
    x = "Month",
    y = "Mean Arrival Delay",
    color = "Carrier"
  ) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

```



