---
title: "A4 - Logical Vectors Assignment"
author: "Saransh Rakshak"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, class.source="bg-success")
```

# Loading Packages

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(dbplyr)
#Load any additional packages
library(nycflights13)

```

# Dataset

```{r}
# Display flights dataset
flights
```


# filter()

<br>

## Flights that had an arrival delay of two or more hours

```{r}
flights %>% filter(arr_delay >= 120)
```



## Flights that Flew to Houston (IAH or HOU)

```{r}
flights %>% filter(dest %in% c("IAH", "HOU"))
```


## Flights that were operated by United, American, or Delta

```{r}
flights %>% filter(carrier %in% c("UA", "AA", "DL"))
```


## Flights that Departed in summer (July, August, and September)

```{r}
flights %>% filter(month %in% c(7, 8, 9))
```

## Flights that arrived more than two hours late, but didn’t leave late

```{r}
flights %>% filter(arr_delay >= 120 & dep_delay <= 0)
```


## Flights that Were delayed by at least an hour, but made up over 30 minutes in flight

```{r}
flights %>% filter(dep_delay >= 60 & arr_delay < dep_delay - 30)
```


