---
title: "A4: Strings and Regular Expressions - Template"
author: "Saransh Rakshak"
output: html_document
---

# Setup

```{r warning=FALSE, message=FALSE}
#load packages
library(tidyverse)
library(dbplyr)
library(stringr)
library(tidyr)
```

__1. Create a regular expression to find (create a subset of) all words that start with "y". __


```{r}
start_with_y <- str_subset(words, "^y")
start_with_y
```

<br>


__2. Create a regular expression to find (create a subset of) all words that end with "x". __

```{r}
end_with_x <- str_subset(words, "x$")
end_with_x
```

<br>


__3. Create a regular expression to find (create a subset of) all words that have exactly 3 letters. __

```{r}
words_3_chars <- str_subset(words, "^.{3}$")
words_3_chars
```

<br>

__4. Create a regular expression to find (create a subset of) all words that have 7 letters or more. __


```{r}
seven_or_more_letters <- str_subset(words, "^.{7,}$")
seven_or_more_letters
```

<br>


__5. Create a regular expression to find (create a subset of) all words that start with a vowel. __
 
```{r}
words_starting_with_vowel <- str_subset(words, "^[aeiouAEIOU]")
words_starting_with_vowel
```

<br>


__6. Create a regular expression to find (create a subset of) all words that end in `ing` or `ise`. Hint: use parentheses `()` to include and/or statement. __

```{r}
words_ending_ing_ise <- str_subset(words, "(ing|ise)$")
words_ending_ing_ise
```

<br>


__7. Create a regular expression to find (create a subset of) all words that start with three consonants (i.e., not vowels).__


```{r}
words_starting_with_3_consonants <- str_subset(words, "^[^aeiouAEIOU]{3}")
words_starting_with_3_consonants
```

<br>


__8. Replace all a's in `words` with an underscore `_`.  (using `head` function only display first 15 observations)__


```{r}
words_with_underscores <- str_replace_all(words, "a", "_")
head(words_with_underscores, 15)
```

<br>

__9. Using the `billboard` tibble from `tidyr` (part of `tidyverse`), create a subset of the songs (tracks) in the top 100 in the year 2000 (all songs in the `billboard` tibble are from the year 2000) that contain the word "Love".__

```{r}
love_songs <- billboard %>% filter(str_detect(track, "Love"))
love_songs
```

<br>

__10. Using the `billboard` tibble, for the song tracks replace all the words "Love" with "LOVE". Only return those tracks that have the word replaced.  (Hint: you may have to do this is two or more steps OR you can use piping)__

```{r}
billboard %>%
  filter(str_detect(track, "Love")) %>%
  mutate(track = str_replace_all(track, "Love", "LOVE"))
```




