---
title: "A3_MoreOnGraphs"
author: "Saransh Rakshak"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, class.source="bg-success")
```

### Setup

```{r message=FALSE, warning=FALSE}
# Load packages needed
library(tidyverse)
library(openintro)
library(visdat)
```


```{r}
# Create condensed dataframe loans
loans <- loans_full_schema %>%
  select(loan_amount, interest_rate, term, grade, 
         state, annual_income, homeownership, debt_to_income)

glimpse(loans)
```

<br>

#### Visual of types of data in `loans` dataframe
```{r}
vis_dat(loans)
```


<br>

#### Using the `loans` dataframe create a histogram for `interest_rate`.  The histogram should have an appropriate amount of bins.  Be sure to provide an apprppriate title and axis labels.

```{r}
ggplot(loans, aes(x = interest_rate)) +
  geom_histogram(binwidth = 5) +
  labs(
    x = "Interest Rate (%)",
    y = "Frequency",
    title = "Amounts of Interest Rate loans"
  )
```


```{r}
ggplot(loans, aes(x = interest_rate)) +
  geom_histogram(binwidth = 10) +
  labs(
    x = "Interest Rate (%)",
    y = "Frequency",
    title = "Amounts of Interest Rate loans"
  )
```


#### Describe the shape of the distribution of `interest_rate`. 

- Based on binwidth = 1 histogram, we can see the distribution of interest_rate to be multimodal and right skewed. With binwidth = 2 histogram, our distribution of interest_rate becomes unimodal but remains right skewed, with the largest distribution of interest_rate being between 8% and 12%.

<br>

#### Using the `loans` dataframe, create a facet histogram of `interest_rate` broken down by `grade`.

```{r}
ggplot(loans, aes(x = interest_rate, fill = grade)) + 
  geom_histogram(binwidth = 5) +
  labs(
    x = "Interest Rate (%)",
    y = "Frequency",
    title = "Amounts of Interest Rate loans"
  ) +
  facet_grid(~grade)
```

<br>

#### Using the `loans` dataframe, create a density plot of `interest_rate` separated by `grade`.  Use an appropriate adjustment value.  Use appropriate titles and labels.  


```{r}
ggplot(loans, aes(x = interest_rate, 
                  fill = grade)) +
  geom_density(adjust = 5, 
               alpha = 0.5) +
  labs(
    x = "Interest Rate (%)",
    y = "Density",
    title = "Amounts of Interest Rate loans", 
    fill = "Grade"
  )
```

#### Do you notice anything unusual about this graph?  Explain. 

- Yes. For grades A-F there is distribution spread across a range of interest rates. But, for grade G, there is little to no spread, this may likely be the result of those in grade G receiving a shared, fixed value interest rate. This is also reflected in our facet histogram above and our boxplot below, having values for only one specific interest rate percentage.

<br>

#### Using the loans dataframe, create boxplots of `interest_rate` sepearted by `grade`, coloring each box a different color based on `grade`.  Use appropriate labels and titles. 

```{r}
ggplot(loans, aes(x = grade,
                  y = interest_rate, fill=grade)) +
  geom_boxplot() +
  labs(
    x = "Grade",
    y = "Interest Rate (%)",
    title = "Amounts of Interest Rate loans", 
    subtitle = "by type of grade"
  )
```



<br>

#### Using the loans dataframe create a hexplot of `loan_amount` vs. `interest_rate` for `grade` (Hint: use the `fill` argument).  Please provide appropriate title(s) and labels. 

```{r}
ggplot(loans, aes(x = loan_amount, y = interest_rate, color = grade), fill = grade) +
  geom_hex() + 
  labs(
    x = "Loan Amount ($)",
    y = "Interest Rate (%)",
    title = "Hexplot of Loan Amount versus Interest Rate, outlined by Grade"
  ) + 
  scale_fill_viridis_c()
```

<br>

#### Using the loans dataframe and the idea that you can filter within the ggplot function, create a hexplot of `loan_amount` vs. `interest_rate` for `grade` values A thru F (exclude grade G).  Please provide appropriate title(s) and labels. 

```{r}
ggplot(loans %>% filter(grade != "G"),
       aes(x = loan_amount, y = interest_rate, color = grade), fill = grade) +
  geom_hex() + 
  labs(
    x = "Loan Amount ($)",
    y = "Interest Rate (%)",
    title = "Hexplot of Loan Amount versus Interest Rate, outlined by Grade"
  ) + 
  scale_fill_viridis_c()
```

<br>


#### Using the `loans` dataframe, create a segmented bar chart of `state` segmented by `homeownership`.  The bars should be horizontal (not vertical) and the length of the bars should be the count for each state.  The graph should have appropriate titles, and labels.

```{r, fig.width=14}
ggplot(loans, aes(x = state, 
                  fill = homeownership)) +
  geom_bar() +
    labs(
    x="State", 
    y= "Count",
    title= "Frequency Chart of State"
  )
```








