---
title: "A3_EDA_Template"
author: "YOUR NAME HERE"
output: 
   html_document:
     df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, class.source="bg-success")
```

### Setup

#### Loading packages
```{r message=FALSE, warning=FALSE}
# Load packages needed
```

#### Loans dataframe
```{r}
# Create condensed dataframe loans

```

<br>

### Univariate Numerical Summary

Using the `summarise` function, create a summary of the following statistics. 

- median of `interest_rate`
- IQR (function `iqr()`) of `interest_rate`
- mean of `debt_to_income`
- standard deviation of `debt_to_income`

```{r}
# univariate numerical summary

```

<br>

### Categorical Summary

Create a frequency table of `grade`. 

```{r}
# frequency table of grade

```


<br>

### Multivariate Numerical Summary

Create a summary statistics table for the following statistics grouped by `grade`

- median of `interest_rate`
- IQR (function `iqr()`) of `interest_rate`
- mean of `debt_to_income`
- standard deviation of `debt_to_income`


```{r}
# Multivariate numerical summary

```

<br>

### Multivariate Categorical Summary

Create a cross table of frequencies for `homeownership` by `grade`.  

```{r}
# Multivariable Categorical Summary 

```

