---
title: "A3_EDA_Template"
author: "Saransh Rakshak"
output: 
   html_document:
     df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, class.source="bg-success")
```

### Setup

#### Loading packages
```{r message=FALSE, warning=FALSE}
# Load packages needed
library(tidyverse)
library(openintro)
```

#### Loans dataframe
```{r}
# Create condensed dataframe loans
loans <- loans_full_schema %>%
  select(loan_amount, interest_rate, term, grade, 
         state, annual_income, homeownership, debt_to_income)
```

<br>

### Univariate Numerical Summary

Using the `summarise` function, create a summary of the following statistics. 

- median of `interest_rate`
- IQR (function `iqr()`) of `interest_rate`
- mean of `debt_to_income`
- standard deviation of `debt_to_income`

```{r}
# univariate numerical summary
loans |>
  summarise(median_interest_rate = median(interest_rate),
            interest_IQR = IQR(interest_rate),
            mean_debt_to_income = mean(debt_to_income, na.rm=TRUE),
            sd_debt_to_income = sd(debt_to_income, na.rm=TRUE))
```

<br>

### Categorical Summary

Create a frequency table of `grade`. 

```{r}
# frequency table of grade
loans |>
  count(grade)
```


<br>

### Multivariate Numerical Summary

Create a summary statistics table for the following statistics grouped by `grade`

- median of `interest_rate`
- IQR (function `iqr()`) of `interest_rate`
- mean of `debt_to_income`
- standard deviation of `debt_to_income`


```{r}
# Multivariate numerical summary
loans |>
  group_by(grade) |>
  summarize(median_interest_rate = median(interest_rate),
            interest_IQR = IQR(interest_rate),
            mean_debt_to_income = mean(debt_to_income, na.rm=TRUE),
            sd_debt_to_income = sd(debt_to_income, na.rm=TRUE))
```

<br>

### Multivariate Categorical Summary

Create a cross table of frequencies for `homeownership` by `grade`.  

```{r}
# Multivariable Categorical Summary 
loans |>
  group_by(homeownership) |>
  count(grade)
```

